---
title: "San Francisco Trees"
output: html_notebook
- "Vicki M. Zhang"
- "TidyTuesdays (Jan 28th, 2020)"
date: "June 1, 2020"
output: 
   html_document:
      toc: true
      toc_depth: 4
      toc_float: true
      df_print: paged
      code_folding: show
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```


# Introduction

I learn more stats!

This [data set](https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv) contains information on tree species, their locations (address, latitude, and longitude), the date it was planted, some of it's measurements, and other variables. Some of the data is missing (obviously). I really like this data set because it's huge, and it's messy.

```{r packages}
library(tidyverse)
library(naniar)
```

 
```{r data}
# sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')
```

# Explore the data

To explore the data, I followed the first parts of Julia Silge's tutorial [Tuning random forest hyperparameters with #TidyTuesday trees data](https://juliasilge.com/blog/sf-trees-random-tuning/).There are a lot of variables, and the data ain't clean.

```{r}
glimpse(sf_trees)
sf_trees
```


## Clean

First, variables. Code all cases where `legal_status` is not "DPW Maintained" as "Other". Then, use `parse_number()` to get a rough estimate of the size of the plot from the `plot_size` column. Next, remove the `address` variable (redundant info as there is also the more accurate lat/long). Finally, change all character variables to factor.

```{r}
trees <- sf_trees %>% 
  mutate(
    
    # mutate so that `legal_status` is either "DPW Maintained" or "Other"
    legal_status = case_when(
      legal_status == "DPW Maintained" ~ legal_status,
      TRUE ~ "Other"
    ),
    
    # parse the `plot_size` to a number
    plot_size = parse_number(plot_size)
  ) %>% 
  
  # remove address
  select(-address) %>% 
  
  
  # character to factor
  mutate_if(is.character, factor)
```
## NAs

Before removing any NAs, let's just check them.

```{r}
# visualize
vis_miss(trees, warn_large_data = FALSE)

# missing data summaries for each variable
miss_var_summary(trees)

# span of missing data by `date` and `plot_size`
miss_var_span(trees, var = date, span_every = 5000)
miss_var_span(trees, var = plot_size, span_every = 5000)
```

About 11% of the data is missing, spread across the 6 out of the 11 variables. The majority of these NAs are missing `date` and `plot_size`. Most of the later `date` values are missing, while most of the earlier `plot_size` values are missing.

Remove all NAs.

```{r}
trees <- trees %>% 
  na.omit()
```

## Exploratory Data Analysis

*How are the trees distributed across San Fran?*

This is cool - plot `latitude` and `longitude` without needing to draw it over a map. Obviously it is missing some attributes (e.g., scale), but it does give a brief visualization and shows obvious streets.

```{r}
trees %>% 
  ggplot(aes(x = longitude, y = latitude, colour = legal_status)) + 
  geom_point(size = 0.5, alpha = 0.5) +
  labs(colour = NULL) + # remove colour legend title
  theme_classic()
```
*What relationships do we see with the caretaker of each tree?*

First, use `count()` to create a contingency table that tabulates the number of counts `n` of caretakers grouped by `legal_status`. Then, since there is already an `n` column, pipe this into `add_count()`. The `add_count()` function is a combination of `group_by()` (here, grouping by caretaker) and `add_tally()` (counting the `n` number of caretakers within each caretaker group). Note that the column `n` is still hanging around. So, `caretaker_count` is the total tally of that group of caretakers, while `n` groups by `caretakers` and `legal_status`. Without using `(wt = n)`, only the number of rows for each caretaker would be counted (i.e., the number of different `legal_status` statuses).

Then, filter only cases where the `caretaker_count` is greater than 50. Group by `legal_status`, and create a `percent_legal` column that that calculates the percentage of caretakers, separated by legal status.

Create a vertical bar graph using the new `percent_legal` column.

```{r}
trees %>% 
  
  # create contingency table of `caretaker` and `legal_status`
  count(legal_status, caretaker) %>% 
  
  # add the weighted count of each caretaker as a column named `caretaker_count`
  add_count(caretaker, wt = n, name = "caretaker_count") %>% 
  
  # filter by cases with caretaker > 50
  filter(caretaker_count > 50) %>% 
  
  # calculate percentage of caretakers for each `legal_status`
  group_by(legal_status) %>% 
  mutate(percent_legal = n / sum(n)) %>% 
  
  # create bar graph
  ggplot(aes(x = percent_legal, y = caretaker, fill = legal_status)) +
  geom_col(position = "dodge") +
  labs(fill = NULL, # remove fill legend title
       x = "% of trees grouped by legal status") +
  theme_classic()
```

From the bar graph, some things can be noticed. First, a lot of caretakers are private, and there are about twice as many private caretakers that maintain trees with the legal status of "other" than the DPW maintained status. Trees in DPW-maintained areas are split between private caretakers and DPW caretakers (naturally).


# 
























